---
title: "Calibrate rsofun with photocold branch"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)
library(ingestr)
library(knitr)
library(lubridate)
library(rbeni)
```


## Load data


```{r message=FALSE, warning=FALSE}
load("data/df_daily.RDA") # reads df_final

ddf <- df_final %>% 
  as_tibble() %>%
  mutate(doy = lubridate::yday(date)) %>% 
  mutate(greenup = ifelse(doy > sos & doy < peak, TRUE, FALSE))
```

```{r}
visdat::vis_miss(ddf, warn_large_data = FALSE)
```

PROBLEM: Lots of temperature and VPD data is missing. Make sure to get complete data. Should be available (P-model outputs are available too).

## Empirical LUE models

Determine bias of early season bias of empirical LUE models by fitting models outside greenup period.

```{r}
ddf <- ddf %>% 
  mutate(lue = gpp_obs / (fapar * ppfd)) %>% 
  mutate(lue = remove_outliers(lue))

## take mean LUE for constant-LUE model
mod_constlue <- ddf %>% 
  filter(!greenup) %>% 
  pull(lue) %>% 
  mean(., na.rm = TRUE)
  
## LUE as a linear function of temp and vpd
mod_lue_temp_vpd <- lm(lue ~ temp + vpd, 
                       data = ddf %>% 
                         filter(!greenup))
ddf <- ddf %>% 
  mutate(lue_temp_vpd = predict(mod_lue_temp_vpd, newdata = .)) %>% 
  mutate(gpp_temp_vpd = lue_temp_vpd * fapar * ppfd,
         gpp_lue_const = mod_constlue * fapar * ppfd)
```

Compare models
```{r}
## P-model
ddf %>% 
  analyse_modobs2("gpp_pmodel", "gpp_obs", type = "hex")

## constant LUE model
ddf %>% 
  analyse_modobs2("gpp_lue_const", "gpp_obs", type = "hex")

## LUE ~ temp + VPD model
ddf %>% 
  analyse_modobs2("gpp_temp_vpd", "gpp_obs", type = "hex")
```

## Mean seasonal cycle

... per site
```{r}
df_meandoy <- ddf %>% 
  group_by(sitename, doy) %>% 
  summarise(across(starts_with("gpp_"), mean, na.rm = TRUE))
```

### Plot by site

```{r, out.height="500%"}
df_meandoy %>% 
  pivot_longer(c(gpp_obs, gpp_pmodel, gpp_bess, gpp_rf, gpp_lue_const, gpp_temp_vpd), names_to = "model", values_to = "gpp") %>% 
  ggplot() +
  # geom_ribbon(
  #   aes(x = doy, ymin = obs_min, ymax = obs_max), 
  #   fill = "black", 
  #   alpha = 0.2
  #   ) +
  geom_line(aes(x = doy, y = gpp, color = model), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~sitename, ncol = 3 ) +    # , labeller = labeller(climatezone = list_rosetta)
  theme_gray() +
  theme(legend.position = "bottom")
  # scale_color_manual(
  #   name="Model: ",
  #   values=c("red", "black", "royalblue", "darkgoldenrod", "springgreen", "orchid4")
  #   )

ggsave("fig/gpp_meandoy.pdf", height = 25, width = 8)
```

## Normalise to peak season

... by site, using the upper 25% quantile.
```{r}
norm_to_peak <- function(df, mod, obs){
  
  q75_obs <- quantile(df[[obs]], probs = 0.75, na.rm = TRUE)
  q75_mod <- quantile(df[[mod]], probs = 0.75, na.rm = TRUE)
  
  ## normalise mod
  df[[mod]] <- df[[mod]] * 
    mean(df[[obs]][df[[obs]]>q75_obs], na.rm = TRUE) / 
    mean(df[[mod]][df[[obs]]>q75_obs], na.rm = TRUE)
  
  return(df)
}

ddf_norm <- ddf %>% 
  group_by(sitename) %>% 
  nest() %>% 
  mutate(data = purrr::map(data, ~norm_to_peak(., "gpp_pmodel", "gpp_obs"))) %>% 
  mutate(data = purrr::map(data, ~norm_to_peak(., "gpp_bess", "gpp_obs"))) %>% 
  mutate(data = purrr::map(data, ~norm_to_peak(., "gpp_rf", "gpp_obs"))) %>% 
  mutate(data = purrr::map(data, ~norm_to_peak(., "gpp_lue_const", "gpp_obs"))) %>% 
  mutate(data = purrr::map(data, ~norm_to_peak(., "gpp_temp_vpd", "gpp_obs"))) %>% 
  unnest(data)
```

### Plot normalised by site 

```{r}
df_meandoy_norm <- ddf_norm %>% 
  group_by(sitename, doy) %>% 
  summarise(across(starts_with("gpp_"), mean, na.rm = TRUE))
```

```{r, out.height="500%"}
df_meandoy_norm %>% 
  pivot_longer(c(gpp_obs, gpp_pmodel, gpp_bess, gpp_rf, gpp_lue_const, gpp_temp_vpd), names_to = "model", values_to = "gpp") %>% 
  ggplot() +
  # geom_ribbon(
  #   aes(x = doy, ymin = obs_min, ymax = obs_max), 
  #   fill = "black", 
  #   alpha = 0.2
  #   ) +
  geom_line(aes(x = doy, y = gpp, color = model), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~sitename, ncol = 3, scales = "free_y" ) +
  theme_gray() +
  theme(legend.position = "bottom")
  # scale_color_manual(
  #   name="Model: ",
  #   values=c("red", "black", "royalblue", "darkgoldenrod", "springgreen", "orchid4")
  #   )

ggsave("fig/gpp_meandoy_norm.pdf", height = 25, width = 8)
```

