---
title: "Calibrate rsofun with photocold branch"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ingestr)
library(knitr)
library(lubridate)
```


```{r message=FALSE, warning=FALSE}
# devtools::install_github("computationales/rsofun@v4.2")
source("~/photocold/R/wrap_calib_photocold.R")
```

Short version using function
```{r}
## Get model forcing
load("~/data/rsofun_benchmarking/df_drivers_fluxnet2015.Rdata")
```


```{r}
df_drivers_fluxnet2015 <- df_drivers_fluxnet2015 %>% 
  dplyr::select(sitename, forcing) %>% 
  unnest(forcing) %>% 
  dplyr::filter(!(month(date)==2 & mday(date)==29)) %>% 
  
  ## model requires flux per seconds now
  mutate(prec = prec / (60*60*24), ppfd = ppfd / (60*60*24)) %>% 
  
  ## assuming all precipitation in liquid form
  mutate(rainf = prec, snowf = 0) %>% 

  ## required for new version, but not used because   
  mutate(tmin = temp, tmax = temp) %>% 
  
  group_by(sitename) %>% 
  nest() %>%
  rename(forcing = data) %>% 
  right_join(
    df_drivers_fluxnet2015 %>% 
      dplyr::select(-forcing),
    by = "sitename"
  ) %>% 
  ungroup() %>% 
  rename(site_info = siteinfo, params_soil = df_soiltexture)

## change name to make compatible
df_drivers_fluxnet2015 <- df_drivers_fluxnet2015 %>% 
  mutate(forcing = purrr::map(forcing, ~rename(., rain = rainf, snow = snowf)))
```

## Calibration target

```{r warning=FALSE, message=FALSE}
filn <- "~/data/rsofun_benchmarking/ddf_fluxnet_gpp.Rdata"
if (!file.exists(filn)){
  settings_ingestr_fluxnet <- list(
    dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", 
    getswc = FALSE,
    filter_ntdt = TRUE,
    threshold_GPP = 0.8,
    remove_neg = FALSE
  )
  
  ddf_fluxnet_gpp <- ingestr::ingest(
    siteinfo = siteinfo_fluxnet2015 %>% 
      dplyr::filter(sitename %in% calibsites),
    source    = "fluxnet",
    getvars = list(gpp = "GPP_NT_VUT_REF",
                   gpp_unc = "GPP_NT_VUT_SE"),
    dir = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
    settings = settings_ingestr_fluxnet,
    timescale = "d"
    )
  save(ddf_fluxnet_gpp, file = filn)
} else {
  load(filn)
}
```

## Calibration settings

```{r}
settings <- list(
  method      = "bayesiantools",
  targetvars  = c("gpp"),
  timescale   = list(targets_obs = "y"),
  metric      = cost_mse_photocold,
  dir_results = "./",
  name        = "photocold",

  control     = list(
    sampler = "DEzs",
    settings = list(
      burnin = 1500,
      iterations = 5000
    )),
  
  par = list(
    kphio       = list(lower=0.03, upper=0.2,  init = 0.09),
    
    soilm_par_a = list(lower=0.00, upper=1.0,  init = 0.33),
    soilm_par_b = list(lower=0.00, upper=10.0, init = 1.5),
    
    kphio_par_a = list(lower=-10,  upper=10,   init = 5.00),
    kphio_par_b = list(lower=0.1,  upper=10.0, init = 0.50),
    
    kphio_par_c = list(lower=1,    upper=200,  init = 50.0),
    kphio_par_d = list(lower=0.01, upper=10,   init = 0.1),
    kphio_par_e = list(lower=0   , upper=10,   init = 5)
  )
)
```


## Calibrate by sites

```{r}
## my own site selection
df_sites <- ingestr::siteinfo_fluxnet2015 %>% 
  filter(sitename %in% df_drivers_fluxnet2015$sitename) %>% 
  dplyr::filter(koeppen_code %in% c("Cfa", "Cfb", "Cfc", "Dfa", "Dfb", "Dfc", "Dfd") &
                classid %in% c("MF", "ENF", "DBF", "DNF"))
vec_sites <- df_sites$sitename

## fit parameters individually per site
df_params <- purrr::map_dfr(
  as.list(vec_sites),
  ~wrap_calib_photocold(
    .,
    df_drivers_fluxnet2015,
    ddf_fluxnet_gpp,
    settings
  ))

saveRDS(df_params, file = "~/photocold/data/df_params_level_site.rds")
```


## Site-level climate metrics

```{r}
## determine sites for which calibration was successful
use_sites <- df_params %>% 
  drop_na() %>% 
  pull(sitename)

## get tmin
ddf_tmin <- ingest(
  siteinfo  = siteinfo_fluxnet2015 %>% filter(sitename %in% use_sites),
  source    = "fluxnet",
  getvars   = list(tmin = "TMIN_F"),
  dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",  # adjust this with your local path
  settings  = list(
    dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", # adjust this with your local path
    dir_hr = "~/data/FLUXNET-2015_Tier1/20191024/HH/", # adjust this with your local path
    getswc = FALSE),
  timescale = "d",
  verbose = TRUE
  )

## add daily mean temperature
df_meteo <- ddf_tmin %>% 
  unnest(data) %>% 
  left_join(
    df_drivers_fluxnet2015 %>% 
      select(sitename, forcing) %>% 
      unnest(forcing) %>% 
      select(-tmin, -tmax),
    by = c("sitename", "date")
  )

## mean seasonal cycle: mean temp, minimum tmin
df_meandoy <- df_meteo %>% 
  mutate(doy = lubridate::yday(date)) %>% 
  group_by(sitename, doy) %>% 
  summarise(temp = mean(temp, na.rm = TRUE),
            tmin = min(tmin, na.rm = TRUE))

## get maximum temperature difference between mean and min seasonal cycles 
df_tdiff <- df_meandoy %>% 
  filter(doy %in% seq(60:150)) %>% 
  mutate(tdiff = temp - tmin) %>% 
  group_by(sitename) %>% 
  summarise(tdiff = mean(tdiff, na.rm = TRUE))
```


## Analyse correlations

```{r}
df <- df_params %>% 
  left_join(
    df_tdiff,
    by = "sitename"
  )

df <- df %>% 
  left_join(
    siteinfo_fluxnet2015,
    by = "sitename"
  )

df$kphio_par_c %>% hist()

df %>% 
  # mutate(classid = ifelse(classid == "MF", "DBF", classid)) %>%  # group mixed and deciduous forests together
  ggplot(aes(tdiff, kphio_par_c, color = classid)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()

linmod_dbf <- lm(kphio_par_c ~ tdiff, 
   data = df %>% 
     # mutate(classid = ifelse(classid == "MF", "DBF", classid)) %>%  # group mixed and deciduous forests together
     filter(classid == "DBF") %>% 
     drop_na() %>% 
     filter(!is.infinite(tdiff))
     )
summary(linmod_dbf)

df %>% 
  mutate(classid = ifelse(classid == "MF", "DBF", classid)) %>%  # group mixed and deciduous forests together
  ggplot(aes(classid, kphio_par_c)) +
  geom_boxplot()

df %>% 
  mutate(classid = ifelse(classid == "MF", "DBF", classid)) %>%  # group mixed and deciduous forests together
  ggplot(aes(classid, kphio_par_a)) +
  geom_boxplot()
```




## Run model

Update model parameters.
```{r}
params_modl <- list(
    kphio       = pars$par[1],
    
    soilm_par_a = pars$par[2],
    soilm_par_b = pars$par[3],
    
    kphio_par_a = pars$par[4],
    kphio_par_b = pars$par[5],
    kphio_par_c = pars$par[6],
    kphio_par_d = pars$par[7],
    kphio_par_e = pars$par[8]
  )
params_modl
```


```{r warning=FALSE, message=FALSE}
output <- rsofun::runread_pmodel_f(
  df_drivers_fluxnet2015 %>% 
      dplyr::filter(sitename %in% calibsites) %>% 
      mutate(forcing = purrr::map(forcing, 
                                  ~mutate(., 
                                          snow = ifelse(temp < 1, prec, 0),
                                          prec = ifelse(temp < 1, 0, prec)))),
  par = params_modl
  )

saveRDS(output, file = "../data/output_photocold.rds")
```

## Run evaluation

### Single site

```{r}
df_modobs <- output %>% 
  filter(sitename %in% calibsites) %>% 
  select(sitename, data) %>% 
  unnest(data) %>% 
  select(sitename, date, mod = gpp) %>% 
  left_join(
    ddf_fluxnet_gpp %>% 
      filter(sitename %in% calibsites) %>% 
      unnest(data) %>% 
      rename(obs = gpp),
    by = c("sitename", "date")
  )
```

```{R}
library(rbeni)
df_modobs %>%
  # ggplot() +
  # geom_hex(aes(mod, obs))
  analyse_modobs2("mod", "obs")

df_modobs %>% 
  ggplot() +
  geom_line(aes(date, obs)) +
  geom_line(aes(date, mod), color = "red")
```

```{R}
df_modobs %>% 
  mutate(doy = lubridate::yday(date)) %>% 
  group_by(sitename, doy) %>% 
  summarise(obs = mean(obs, na.rm = TRUE),
            mod = mean(mod, na.rm = TRUE)) %>% 
  ggplot() +
  geom_line(aes(doy, obs)) +
  geom_line(aes(doy, mod), color = "red")
```

### Multi-site

Do evaluation only for sites where simulation was run.
```{r}
evalsites <- output %>% 
  mutate(ntsteps = purrr::map_dbl(data, ~nrow(.))) %>% 
  dplyr::filter(ntsteps > 0) %>% 
  pull(sitename)
```

Load standard benchmarking file with observational data for evaluation.
```{r}
load("~/data/rsofun_benchmarking/obs_eval_fluxnet2015.Rdata")
```

Define evaluation settings.
```{r}
settings_eval <- list(
  benchmark = list( gpp = c("fluxnet") ),
  sitenames = evalsites,
  agg       = 8
  )
```

And finally run the evaluation.
```{r warning=FALSE, message=FALSE, error=FALSE}
source("~/sofunCalVal/R/eval_sofun.R")
source("~/sofunCalVal/R/get_stats.R")
filn <- "../out_eval_photocold.rds"
overwrite <- TRUE
if (!file.exists(filn) || overwrite){
  out_eval <- eval_sofun( 
    output, 
    settings_eval, 
    settings_sims, 
    obs_eval = obs_eval, 
    overwrite = TRUE, 
    light = FALSE 
    )
  saveRDS(out_eval, file = filn)
} else {
  out_eval <- read_rds(filn)
}
```

## Evaluation results

### Metrics table

```{r}
out_eval$gpp$fluxnet$metrics %>% 
  bind_rows(.id = "Level") %>% 
  kable
```

### Visualisations

#### Correlations

```{r message=FALSE, warning=FALSE}
out_eval$gpp$fluxnet$plot$gg_modobs_xdaily
out_eval$gpp$fluxnet$plot$gg_modobs_spatial_annual
```

#### Mean seasonal cycle

```{r message=FALSE, warning=FALSE}
## plot
out_eval$gpp$fluxnet$data$meandoydf_byclim %>% 
  dplyr::filter(climatezone %in% c("Aw south", "BSk north", "Cfa north", "Cfb north", "Cfb south", "Csa north", "Csb north", "Dfb north", "Dfc north")) %>%
  dplyr::filter(koeppen_code != "-") %>% 
  pivot_longer(c(obs_mean, mod_mean), names_to = "source", values_to = "gpp") %>% 
  ggplot() +
  geom_ribbon(
    aes(x = doy, ymin = obs_min, ymax = obs_max), 
    fill = "black", 
    alpha = 0.2
    ) +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +    # , labeller = labeller(climatezone = list_rosetta)
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name="Setup: ",
    values=c("red", "black")
    # values=c("FULL" = "#DE1A1A", "Observed" = "black")
    )
```

<<<<<<< HEAD
#### Drought response

```{r message=FALSE, warning=FALSE}
source("../R/align_events.R")
source("../R/eval_droughtresponse.R")

df_dday_agg <- eval_droughtresponse( 
  df = out_eval$gpp$fluxnet$data$ddf %>% rename(site = sitename), 
  path_flue = "~/data/flue/flue_stocker18nphyt.csv",
  before=20, 
  after=105,
  leng_threshold = 10, 
  nbins=10, 
  do_norm=TRUE
  )

usecol = colorRampPalette( c("wheat3", "white") )( 5 )[2]

df_dday_agg %>% 
  ggplot() +
  # annotate("rect", xmin = 0, xmax = 105, ymin = -1.2, ymax = 2.2, alpha = .8, fill = usecol) +
  geom_hline(yintercept = 0, color = "black", linetype = "dotted") +
  geom_vline(xintercept = 0, color = "black", linetype = "dotted") +
  geom_line(
    aes(x = dday, y = median), size = 0.9) +
  geom_ribbon(
    aes(x = dday, ymin = q33, ymax = q66), 
    alpha = 0.3) +
  scale_color_manual(values = c("BRC" = "black", "FULL" = "royalblue"), name = "Setup") +
  scale_fill_manual(values = c("BRC" = "black", "FULL" = "royalblue"), name = "Setup") +
  ylim(-1.2, 2.2) + xlim(-20, 105) +
  scale_x_continuous(expand = c(0,0)) + scale_y_continuous(expand = c(0,0)) +
  labs(x = "Days after drought onset", y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) ) +
  theme_classic()
```

## Appendix

### Site list

```{r}
siteinfo_fluxnet2015 %>% 
  dplyr::filter(sitename %in% evalsites) %>% 
  kable()
```
=======
<!-- #### Drought response -->

<!-- ```{r message=FALSE, warning=FALSE} -->
<!-- source("../R/align_events.R") -->
<!-- source("../R/eval_droughtresponse.R") -->

<!-- df_dday_agg <- eval_droughtresponse(  -->
<!--   df = out_eval$gpp$fluxnet$data$ddf %>% rename(site = sitename),  -->
<!--   path_flue = "~/data/flue/flue_stocker18nphyt.csv", -->
<!--   before=20,  -->
<!--   after=105, -->
<!--   leng_threshold = 10,  -->
<!--   nbins=10,  -->
<!--   do_norm=TRUE -->
<!--   ) -->

<!-- usecol = colorRampPalette( c("wheat3", "white") )( 5 )[2] -->

<!-- df_dday_agg %>%  -->
<!--   ggplot() + -->
<!--   # annotate("rect", xmin = 0, xmax = 105, ymin = -1.2, ymax = 2.2, alpha = .8, fill = usecol) + -->
<!--   geom_hline(yintercept = 0, color = "black", linetype = "dotted") + -->
<!--   geom_vline(xintercept = 0, color = "black", linetype = "dotted") + -->
<!--   geom_line( -->
<!--     aes(x = dday, y = median), size = 0.9) + -->
<!--   geom_ribbon( -->
<!--     aes(x = dday, ymin = q33, ymax = q66),  -->
<!--     alpha = 0.3) + -->
<!--   scale_color_manual(values = c("BRC" = "black", "FULL" = "royalblue"), name = "Setup") + -->
<!--   scale_fill_manual(values = c("BRC" = "black", "FULL" = "royalblue"), name = "Setup") + -->
<!--   ylim(-1.2, 2.2) + xlim(-20, 105) + -->
<!--   scale_x_continuous(expand = c(0,0)) + scale_y_continuous(expand = c(0,0)) + -->
<!--   labs(x = "Days after drought onset", y = expression( paste( "Bias (g C m"^{-1}, " d"^{-1}, ")")) ) + -->
<!--   theme_classic() -->
<!-- ``` -->

<!-- ## Appendix -->

<!-- ### Site list -->

<!-- ```{r} -->
<!-- siteinfo_fluxnet2015 %>%  -->
<!--   dplyr::filter(sitename %in% evalsites) %>%  -->
<!--   kable() -->
<!-- ``` -->
>>>>>>> 8891d003fd1a53d8d57568db1ffd0ccf779546e7
