---
title: "Explore stress"
author: "Beni"
date: "11/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Load data

Load modelled and observed from global calibration at tag v4.2. Subset to required sites only.

```{r cars}
out_eval <- read_rds("~/data/rsofun_benchmarking/out_eval_v4.2.rds")

## add modelled and observed to dates and sites required for this here
ddf <- read_rds("../data/df_output_rsofun.rds") %>% 
  dplyr::select(sitename, data) %>% 
  unnest(data) %>% 
  dplyr::select(sitename, date) %>% 
  left_join(
    out_eval$gpp$fluxnet$data$ddf,
    by = c("sitename", "date")
  )

# df_sites <- read_rds("data/df_sites_modis_era.csv") %>% 
#   dplyr::select(sitename, lat, koeppen_code)

# drivers
ddf_drivers <- read_rds("../data/p_model_fluxnet_drivers.rds") %>% 
  dplyr::select(sitename, forcing) %>% 
  unnest(forcing) %>% 
  mutate(ppfd = 1e6 * ppfd)

# # observed gpp
# ddf_obs <- read_rds("data/ddf_fluxnet_gpp.rds") %>% 
#   unnest(data)

## combine
ddf <- ddf %>% 
  
  ## add ppfd
  left_join(
    ddf_drivers %>% 
      dplyr::select(sitename, date, ppfd, tmin, temp),
    by = c("sitename", "date")
  )
  
  # ## add observed gpp
  # left_join(
  #   ddf_obs,
  #   by = c("sitename", "date")
  # )
  
  # left_join(
  #   df_sites,
  #   by = "sitename"
  # ) %>% 
  # rename(obs = gpp) %>% 
  # dplyr::select(-gpp_unc)
```

Some missing data
```{r}
ddf %>% 
  visdat::vis_miss(cluster = FALSE, warn_large_data = FALSE)
```

## Seasonal cycle before correction

```{r}
doydf <- ddf %>% 
  mutate(doy = lubridate::yday(date),
         hemisphere = ifelse(lat > 0, "north", "south")) %>% 
  mutate(climatezone = paste(koeppen_code, hemisphere)) %>% 
  group_by(climatezone, doy) %>% 
  summarise(mod = mean(mod, na.rm = TRUE), obs = mean(obs, na.rm = TRUE)) %>% 
  dplyr::filter(climatezone != "NA NA")
```

```{r}
doydf %>% 
  pivot_longer(c(obs, mod), names_to = "source", values_to = "gpp") %>% 
  ggplot() +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  facet_wrap( ~climatezone ) +    # , labeller = labeller(climatezone = list_rosetta)
  theme_gray() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name="Setup: ",
    values=c("red", "black")
    # values=c("FULL" = "#DE1A1A", "Observed" = "black")
    )
```

## Hardening function

$$
h = \frac{1}{1 + e^{a + bx}}
$$

```{r}
f_hardening <- function(temp, param){
  
  xx <- (-1) * temp # * ppfd
  xx <- param["b"] * (xx + param["a"])
  yy <- 1 / (1 + exp(xx))
  return(yy)  
}

par <- readRDS("../data/pars_calib_photocold.rds")
ggplot() +
  geom_function(fun = f_hardening, 
                args = list(param = c(
                  # "a" = par$par[["kphio_par_a"]], 
                  # "b" = par$par[["kphio_par_b"]]
                  "a" = 5, 
                  "b" = 0.5
                  ))) +
  xlim(-20, 30)
```

## Dehardening function

Has the same functional form as the hardening function, but uses cumulative degree days as an independent variable.
```{r}
f_dehardening <- function(temp, param){
  
  xx <- (-1) * temp # * ppfd
  xx <- param["b"] * (xx + param["a"])
  yy <- 1 / (1 + exp(xx))
  return(yy)  
}

ggplot() +
  geom_function(
    fun = f_dehardening, 
    args = list(param = c(
      "a" = 50, 
      "b" = 0.1
      ))) +
    #   fun = f_dehardening, 
    # args = list(param = c(
    #   "a" = par$par[["kphio_par_c"]], 
    #   "b" = par$par[["kphio_par_d"]]
      # ))) +
  
  xlim(-10, 100)
```


## Algorithm

Example
```{r}
df <- ddf %>% 
  filter(sitename == "US-Ha1")

df %>% 
  ggplot() +
  geom_line(aes(date, temp)) +
  geom_line(aes(date, tmin), color = "royalblue")
```

Applying the hardening/dehardening to the time series.
```{r}
param_harden <- c("a" = 5, "b" = 0.5)
param_deharden <- c("a" = 50, "b" = 0.1)
level_hard <- 1.0  # start without hardening
gdd <- 0
df$f_stress <- rep(NA, nrow(df))

for (idx in seq(nrow(df))){
  
  ## determine hardening level
  level_hard_new <-  f_hardening(df$tmin[idx], param_harden)
  
  if (level_hard_new < level_hard){
    
    ## entering deeper hardening
    level_hard <- level_hard_new
    
    ## re-start recovery
    gdd <- 0

    # print(paste("Hardening to", level_hard, "on", df$date[idx]))
  }
  
  ## accumulate growing degree days (GDD)
  gdd <- gdd + max(0, (df$temp[idx] - 5.0))
  
  ## de-harden based on GDD. f_stress = 1: no stress
  level_hard <- level_hard + (1-level_hard) * f_dehardening(gdd, param_deharden)
  df$f_stress[idx] <- level_hard
}
```

Run a second time, now with slower recovery
```{r}
param_deharden <- c("a" = 80, "b" = 0.1)
level_hard <- 1.0  # start without hardening
gdd <- 0
df$f_stress_slow <- rep(NA, nrow(df))

for (idx in seq(nrow(df))){
  
  ## determine hardening level
  level_hard_new <-  f_hardening(df$tmin[idx], param_harden)
  
  if (level_hard_new < level_hard){
    
    ## entering deeper hardening
    level_hard <- level_hard_new
    
    ## re-start recovery
    gdd <- 0

    # print(paste("Hardening to", level_hard, "on", df$date[idx]))
  }
  
  ## accumulate growing degree days (GDD)
  gdd <- gdd + max(0, (df$temp[idx] - 5.0))
  
  ## de-harden based on GDD. f_stress = 1: no stress
  level_hard <- level_hard + (1-level_hard) * f_dehardening(gdd, param_deharden)
  df$f_stress_slow[idx] <- level_hard
}
```


Hardening stress function visualised for the first 200 days.
```{r}
df %>%
  slice(1:200) %>% 
  ggplot() +
  geom_line(aes(date, f_stress), color = "red") +
  geom_line(aes(date, f_stress_slow), color = "royalblue")
```

Better fit of multi-year mean seasonal cycle for US-Ha1? Yes.
```{r}
df <- df %>% 
  mutate(mod2 = f_stress * mod)

df %>% 
  mutate(doy = lubridate::yday(date)) %>% 
  group_by(doy) %>% 
  summarise(mod = mean(mod, na.rm = TRUE), mod2 = mean(mod2, na.rm = TRUE), obs = mean(obs, na.rm = TRUE)) %>% 
  pivot_longer(c(obs, mod, mod2), names_to = "source", values_to = "gpp") %>% 
  ggplot() +
  geom_line(aes(x = doy, y = gpp, color = source), size = 0.4) +
  labs(y = expression( paste("Simulated GPP (g C m"^-2, " d"^-1, ")" ) ), 
       x = "DOY") +
  theme(legend.position = "bottom") +
  scale_color_manual(
    name="Setup: ",
    values=c("red", "springgreen3", "black")
    # values=c("FULL" = "#DE1A1A", "Observed" = "black")
    )

```




