---
title: "Forcing data"
author: "Beni"
date: "11/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ingestr)
library(tidyverse)
library(knitr)
library(ingestr)
```

## Read sites

```{r warning=FALSE, message=FALSE}
df_sites <- read_csv(file = "../data/df_sites.csv")
```

## FLUXNET data

```{r eval = FALSE}
# grab fluxnet data ----
df_fluxnet <-
  ingestr::ingest(
    siteinfo  = df_sites,
    source    = "fluxnet",
    getvars   = list(
      temp = "TA_F_DAY",
      prec = "P_F",
      vpd  = "VPD_F_DAY",
      ppfd = "SW_IN_F",
      patm = "PA_F",
      tmin = "TMIN_F"),
    dir       = "~/data/FLUXNET-2015_Tier1/20191024/DD/",
    settings  = list(
      dir_hh = "~/data/FLUXNET-2015_Tier1/20191024/HH/", getswc = FALSE),
    timescale = "d"
      )
```

## CRU data

```{r eval = FALSE}
df_cru <- ingestr::ingest(
  siteinfo  = df_sites,
  source    = "cru",
  getvars   = "ccov",
  dir       = "~/data/cru/ts_4.01/"
  )

saveRDS(df_cru, file = "../data/df_cru.rds")

df_meteo <- df_fluxnet %>%
  tidyr::unnest(data) %>%
  left_join(
    df_cru %>%
      tidyr::unnest(data),
    by = c("sitename", "date")
  ) %>%
  group_by(sitename) %>%
  tidyr::nest()
```

## MODIS

```{r eval = FALSE}
settings_modis <- get_settings_modis(
  bundle            = "modis_fpar",
  data_path         = "~/data/modis_subsets/",
  method_interpol   = "loess",
  network = c("fluxnet", "icos"),
  keep              = TRUE,
  overwrite_raw     = FALSE,
  overwrite_interpol= TRUE,
  n_focal           = 0
  )

df_modis_fpar <- ingest(
  df_sites,
  source = "modis",
  settings = settings_modis,
  parallel = FALSE,
  ncores = 1
  )

## renaming the variable
df_modis_fpar <- df_modis_fpar %>%
  mutate(
    data = purrr::map(data, ~rename(., fapar = modisvar_filled))
    )

saveRDS(df_modis_fpar, file = "../data/df_modis_fpar.rds")
```

## CO2

```{r eval = FALSE}
# . grab CO2 data ----
df_co2 <- ingestr::ingest(
  fluxnet_sites,
  source  = "co2_mlo",
  verbose = FALSE
  )
```

## Model-related stuff

A final set of required ancillary data covers the soil texture properties.

```{r eval = FALSE}
# . set soil parameters ----
df_soiltexture <- bind_rows(
  top    = tibble(
    layer = "top",
    fsand = 0.4,
    fclay = 0.3,
    forg = 0.1,
    fgravel = 0.1
    ),
  bottom = tibble(
    layer = "bottom",
    fsand = 0.4,
    fclay = 0.3,
    forg = 0.1,
    fgravel = 0.1)
)
```

## Simulation parameters

With all data downloads done we have to set some simulation parameters which
will be merged with all driver data. This data includes the number of spin-up
years, and how soil and temperature stress are dealt with.

```{r eval = FALSE}
params_siml <- list(
  spinup             = TRUE,
  spinupyears        = 10,
  recycle            = 1,
  soilmstress        = TRUE,
  tempstress         = TRUE,
  calc_aet_fapar_vpd = FALSE,
  in_ppfd            = TRUE,
  in_netrad          = FALSE,
  outdt              = 1,
  ltre               = FALSE,
  ltne               = FALSE,
  ltrd               = FALSE,
  ltnd               = FALSE,
  lgr3               = TRUE,
  lgn3               = FALSE,
  lgr4               = FALSE
	)
```

With all pieces in place we can use the `rsofun` collect_drivers_sofun()
function to merge all datasets into one nested dataframe, which serves as
input to the `rsofun` p-model. This concludes the formatting of the driver data

```{r eval = FALSE}
p_model_fluxnet_drivers <- rsofun::collect_drivers_sofun(
  site_info      = df_sites,
  params_siml    = params_siml,
  meteo          = df_meteo,
  fapar          = df_modis_fpar,
  co2            = df_co2,
  params_soil    = df_soiltexture
  )
```

## Model run

```{r}
# run the model for these parameters
# optimized parameters from previous work (Stocker et al., 2020 GMD)
params_modl <- list(
    kphio           = 0.09423773,
    soilm_par_a     = 0.33349283,
    soilm_par_b     = 1.45602286,
    tau_acclim_tempstress = 10,
    par_shape_tempstress  = 0.0
  )

output <- rsofun::runread_pmodel_f(
  p_model_fluxnet_drivers,
  par = params_modl
  )

output$data[[1]] %>% 
  ggplot(aes(date, gpp)) +
  geom_line()
```
